### 2020-10-26 journal/_region_.tex
Add Journal folder for extended publication

\message{ !name(paper.tex)}% Created 2020-10-20 mar. 17:19
% Intended LaTeX compiler: pdflatex
\documentclass[Afour,sageh,times]{sagej}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage[most]{tcolorbox}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{siunitx}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
\usepackage{amsmath,amssymb,amsfonts, cases}
\usepackage{algorithmic, graphicx, textcomp}
\usepackage{xcolor, import, hyperref}
\usepackage{subcaption}
\usepackage[USenglish, english]{babel}
\usepackage{tikz}
\usetikzlibrary{shapes.misc,arrows,arrows.meta}
\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\def\volumeyear{2016}
\author{Dehaeze Thomas\affilnum{1} and Collette Christophe\affilnum{2}}
\affiliation{\affilnum{1}Precision Mechatronics Laboratory, University of Liege, Belgium\\
\affilnum{2}BEAMS Department, Free University of Brussels, Belgium\\
\affilnum{3}European Synchrotron Radiation Facility, Grenoble, France}
\email{dehaeze.thomas@gmail.com}
\bibliographystyle{SageH}
\date{}
\title{Active Damping of Rotating Platforms using Integral Force Feedback}
\input{abstract.tex}
\begin{document}

\message{ !name(paper.tex) !offset(-4) }
\input{abstract.tex}
\message{ !name(paper.tex) !offset(511) }

\end{document}