### 2021-02-20 matlab/preamble.tex

\usepackage{float}

\usepackage{caption,tabularx,booktabs}

\usepackage{biblatex}

\usepackage{fontawesome}

\usepackage{caption}
\usepackage{subcaption}

\captionsetup[figure]{labelfont=bf}
\captionsetup[subfigure]{labelfont=bf}
\captionsetup[listing]{labelfont=bf}
\captionsetup[table]{labelfont=bf}

\usepackage{xcolor}

\definecolor{my-blue}{HTML}{6b7adb}
\definecolor{my-pale-blue}{HTML}{e6e9f9}
\definecolor{my-red}{HTML}{db6b6b}
\definecolor{my-pale-red}{HTML}{f9e6e6}
\definecolor{my-green}{HTML}{6bdbb6}
\definecolor{my-pale-green}{HTML}{e6f9f3}
\definecolor{my-yellow}{HTML}{dbd26b}
\definecolor{my-pale-yellow}{HTML}{f9f7e6}
\definecolor{my-orange}{HTML}{dba76b}
\definecolor{my-pale-orange}{HTML}{f9f0e6}
\definecolor{my-grey}{HTML}{a3a3a3}
\definecolor{my-pale-grey}{HTML}{f0f0f0}
\definecolor{my-turq}{HTML}{6bc7db}
\definecolor{my-pale-turq}{HTML}{e6f6f9}

\usepackage{inconsolata}

\usepackage[newfloat=true, chapter]{minted}
\usemintedstyle{autumn}

\setminted{frame=lines,breaklines=true,tabsize=4,fontsize=\scriptsize,autogobble=true,labelposition=topline,bgcolor=my-pale-grey}
\setminted[matlab]{label=Matlab}
\setminted[latex]{label=LaTeX}
\setminted[bash]{label=Bash}
\setminted[python]{label=Python}
\setminted[text]{label=Results}
\setminted[md]{label=Org Mode}

\setmintedinline{fontsize=\normalsize,bgcolor=my-pale-grey}

\usepackage[most]{tcolorbox}

\tcbuselibrary{minted}

\newtcolorbox{seealso}{   enhanced,breakable,colback=my-pale-grey,colframe=my-grey,fonttitle=\bfseries,title=See Also}
\newtcolorbox{hint}{      enhanced,breakable,colback=my-pale-grey,colframe=my-grey,fonttitle=\bfseries,title=Hint}
\newtcolorbox{definition}{enhanced,breakable,colback=my-pale-red, colframe=my-red, fonttitle=\bfseries,title=Definition}
\newtcolorbox{important}{ enhanced,breakable,colback=my-pale-red, colframe=my-red, fonttitle=\bfseries,title=Important}
\newtcolorbox{exampl}[1][]{    enhanced,breakable,colback=my-pale-green,colframe=my-green,fonttitle=\bfseries,title=Example,#1}
\newtcolorbox{exercice}{  enhanced,breakable,colback=my-pale-yellow,colframe=my-yellow,fonttitle=\bfseries,title=Exercice}
\newtcolorbox{question}{  enhanced,breakable,colback=my-pale-yellow,colframe=my-yellow,fonttitle=\bfseries,title=Question}
\newtcolorbox{answer}{    enhanced,breakable,colback=my-pale-turq,colframe=my-turq,fonttitle=\bfseries,title=Answer}
\newtcolorbox{summary}{   enhanced,breakable,colback=my-pale-blue,colframe=my-blue,fonttitle=\bfseries,title=Summary}
\newtcolorbox{note}{      enhanced,breakable,colback=my-pale-blue,colframe=my-blue,fonttitle=\bfseries,title=Note}
\newtcolorbox{caution}{   enhanced,breakable,colback=my-pale-orange,colframe=my-orange,fonttitle=\bfseries,title=Caution}
\newtcolorbox{warning}{   enhanced,breakable,colback=my-pale-orange,colframe=my-orange,fonttitle=\bfseries,title=Warning}

\newtcolorbox{my-quote}[1]{%
  colback=my-pale-grey,
  grow to right by=-10mm,
  grow to left by=-10mm,
  boxrule=0pt,
  boxsep=0pt,
  breakable,
  enhanced jigsaw,
  borderline west={4pt}{0pt}{my-grey}}

\renewenvironment{quote}{\begin{my-quote}}{\end{my-quote}}

\newtcolorbox{my-verse}[1]{%
  colback=my-pale-grey,
  grow to right by=-10mm,
  grow to left by=-10mm,
  boxrule=0pt,
  boxsep=0pt,
  breakable,
  enhanced jigsaw,
  borderline west={4pt}{0pt}{my-grey}}

\renewenvironment{verse}{\begin{my-verse}}{\end{my-verse}}

\usepackage{environ}% http://ctan.org/pkg/environ
\NewEnviron{aside}{%
  \marginpar{\BODY}
}

\renewenvironment{verbatim}{\VerbatimEnvironment\begin{minted}[]{text}}{\end{minted}}

\usepackage{soul}
\sethlcolor{my-pale-grey}

\let\OldTexttt\texttt
\renewcommand{\texttt}[1]{{\ttfamily\hl{\mbox{\,#1\,}}}}

\makeatletter
\preto\Gin@extensions{png,}
\DeclareGraphicsRule{.png}{pdf}{.pdf}{\noexpand\Gin@base.pdf}
\makeatother

\usepackage{hyperref}
\hypersetup{
  colorlinks = true,
  allcolors = my-blue
}

\usepackage{hypcap}